\thetitle

### 2008-05-28 paper.tex
Fixed up figure sizes and some spacing issues.

\newcommand{\thetitle}{An Algebraic Representation of Wireless Workload}
\newcommand{\theauthors}{%
  Stefan~Karpinski \and
  Elizabeth~M.~Belding \and
  Kevin~C.~Almeroth \and
  John~R.~Gilbert%
}

@@

The {packets-bytes-duration} representation is clearly overly simplistic and very incomplete. It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time and flow size unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets, etc. into a vector space. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take a highly detailed but intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the flow type, source node, destination node, start time, flow size, and empirical distributions of packet sizes and inter-packet intervals. If there are $f$ flows in the entire workload pattern, then the collective behavior can be described with six matrices; the rows of each matrix are the behavior vectors of the flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors:\vspace{-0.25em}
\begin{tabbing}
\hspace{1.5em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\parensx{\vec{d}_i}\:$\=$=\parensx{d_{ij}}\:$\=\kill
\>\bullet\> type: \>$T$\>$=\parensx{\vec{t}_i}$\>$=\parensx{t_{ij}}$\>$\in\R^{f \times 255}$,\\
\>\bullet\> sources: \>$S$\>$=\parensx{\vec{s}_i}$\>$=\parensx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\parensx{\vec{d}_i}$\>$=\parensx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> start times: \>$A$\>$=\parensx{\vec{a}_i}$\>$=\parensx{a_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> sizes in bytes: \>$B$\>$=\parensx{\vec{b}_i}$\>$=\parensx{b_{ij}}$\>$\in\R^{f \times d_b}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\parensx{\vec{z}_i}$\>$=\parensx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\parensx{\vec{v}_i}$\>$=\parensx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}
The total behavior is the direct product of these components: $F = \seq{T,S,D,A,B,Z,V} \in \R^{f \times \Sigma}$ where $\Sigma=256+2n+d_t+d_b+d_z+d_v$. Many common traffic models can be expressed succinctly and precisely in terms of transformations or conditions on this behavior matrix and its constituents. In the following sections, we explain in detail how each flow's behavior is represented in these matrices. Finally, we describe how a model instance stochastically generates network traffic.

@@

\begin{figure*}[tb]
\begin{center}
\subfloat[trace node behavior]
{\includegraphics[width=1.392in]{flow_topology_trace}\label{fig:flow-topology-trace}}
\subfloat[uniform node behavior]
{\includegraphics[width=1.392in]{flow_topology_uniform}\label{fig:flow-topology-uniform}}
\subfloat[trace flow behavior]
{\includegraphics[width=1.392in]{flow_behavior_trace}\label{fig:flow-behavior-trace}}
\subfloat[uniform flow behavior]
{\includegraphics[width=1.392in]{flow_behavior_uniform}\label{fig:flow-behavior-uniform}}
\subfloat[packet behavior]
{\includegraphics[width=1.392in]{flow_packet_behavior}\label{fig:packet-behavior}}
\caption{Examples illustrating trace (real) versus uniform (na\"ive synthetic) behaviors at the node, flow and packet levels. Figures \ref{fig:flow-topology-trace} and \ref{fig:flow-topology-uniform} show example node behaviors. The width of each line is proportional to the logarithm of the number of flows between the nodes it connects (zero is the Internet gateway). Uniform and trace flow behavior examples are plotted in Figures \ref{fig:flow-behavior-trace} and \ref{fig:flow-behavior-uniform}. The time axis indicates when flows start and end; the width of each flow line is proportional to the logarithm of its data rate. Figure \ref{fig:packet-behavior} compares uniform (i.e. {\footnotesize{CBR}}) packet behavior with the trace of an actual flow. In the uniform model, the cumulative data sent increases smoothly over time, whereas in the actual packet trace, the transmissions are variable both in size and in inter-packet interval, leading to a ``lumpy'' cumulative data plot.}
\label{fig:trace-vs-uniform}
\end{center}
\vspace{-1.25em}
\end{figure*}

@@

\begin{sidewaysfigure*}
%\begin{figure*}[htb]
\subfloat[\textbf{Application:} Average End-to-End Delay]%
{\includegraphics[width=4.6in]{plots/Application/delay.pdf}}
\subfloat[\textbf{Application:} Packet Delivery Ratio]%
{\includegraphics[width=4.6in]{plots/Application/delivery_ratio.pdf}}
\subfloat[\textbf{Application:} Average Received Throughput]%
{\includegraphics[width=4.6in]{plots/Application/throughput.pdf}}
\subfloat[\textbf{Network:} AODV Control Overhead]%
{\includegraphics[width=4.6in]{plots/Network/control_packets.pdf}}
\subfloat[\textbf{Network:} Packets Dropped in Routing Queue]%
{\includegraphics[width=4.6in]{plots/Network/dropped.pdf}}
\subfloat[\textbf{Link:} 802.11 Control Overhead]%
{\includegraphics[width=4.6in]{plots/Link/overhead.pdf}}%
\caption{Box-and-whisker plots of log-ratio error values for all metrics and traffic models. The lower axis indicates the log-ratio, while the upper axis shows raw ratio values. Each box contains the central majority of log-ratio values: the left and right bounds are at the $25^\text{th}$ and $75^\text{th}$ percentiles. The dark middle line indicates the median value, while the diamond marks the mean. The whiskers (dotted lines) extend to the furthest non-outlier values, while the points beyond that are outliers. The notch in the middle of each bar indicates a $95\%$ confidence interval for the true underlying median value; if two notches do not overlap, they are very unlikely to have the same median.}
\label{fig:box-plots}
%\vspace{-1.25em}
%\end{figure*}
\end{sidewaysfigure*}

>>>>

\newcommand{\thetitle}{Linear Representation of Wireless Workload}
\newcommand{\theauthors}{%
  Stefan~Karpinski \and
  Elizabeth~M.~Belding \and
  Kevin~C.~Almeroth \and
  John~R.~Gilbert%
}

@@

The {packets-bytes-duration} representation is clearly overly simplistic and very incomplete. It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time and flow size unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets, etc. into a vector space. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take a highly detailed but intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the flow type, source node, destination node, start time, flow size, and empirical distributions of packet sizes and inter-packet intervals. If there are $f$ flows in the entire workload pattern, then the collective behavior can be described with six matrices; the rows of each matrix are the behavior vectors of the flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors:\vspace{-0.25em}
\begin{tabbing}
\hspace{1em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\parensx{\vec{d}_i}\:$\=$=\parensx{d_{ij}}\:$\=\kill
\>\bullet\> type: \>$T$\>$=\parensx{\vec{t}_i}$\>$=\parensx{t_{ij}}$\>$\in\R^{f \times 255}$,\\
\>\bullet\> sources: \>$S$\>$=\parensx{\vec{s}_i}$\>$=\parensx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\parensx{\vec{d}_i}$\>$=\parensx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> start times: \>$A$\>$=\parensx{\vec{a}_i}$\>$=\parensx{a_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> sizes in bytes: \>$B$\>$=\parensx{\vec{b}_i}$\>$=\parensx{b_{ij}}$\>$\in\R^{f \times d_b}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\parensx{\vec{z}_i}$\>$=\parensx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\parensx{\vec{v}_i}$\>$=\parensx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}
The total behavior is the direct product of these components: $F = \seq{T,S,D,A,B,Z,V} \in \R^{f \times \Sigma}$ where $\Sigma=256+2n+d_t+d_b+d_z+d_v$. Many common traffic models can be expressed succinctly and precisely in terms of transformations or conditions on this behavior matrix and its constituents. In the following sections, we explain in detail how each flow's behavior is represented in these matrices. Finally, we describe how a model instance stochastically generates network traffic.

@@

\begin{figure*}[tb]
\begin{center}
\subfloat[node trace]
{\includegraphics[width=1.335in]{flow_topology_trace}\label{fig:flow-topology-trace}}
\subfloat[node uniform]
{\includegraphics[width=1.335in]{flow_topology_uniform}\label{fig:flow-topology-uniform}}
\subfloat[flow trace]
{\includegraphics[width=1.335in]{flow_behavior_trace}\label{fig:flow-behavior-trace}}
\subfloat[flow uniform]
{\includegraphics[width=1.335in]{flow_behavior_uniform}\label{fig:flow-behavior-uniform}}
\subfloat[packet]
{\includegraphics[width=1.335in]{flow_packet_behavior}\label{fig:packet-behavior}}
\caption{Examples illustrating trace (real) versus uniform (na\"ive synthetic) behaviors at the node, flow and packet levels. Figures \ref{fig:flow-topology-trace} and \ref{fig:flow-topology-uniform} show example node behaviors. The width of each line is proportional to the logarithm of the number of flows between the nodes it connects (zero is the Internet gateway). Uniform and trace flow behavior examples are plotted in Figures \ref{fig:flow-behavior-trace} and \ref{fig:flow-behavior-uniform}. The time axis indicates when flows start and end; the width of each flow line is proportional to the logarithm of its data rate. Figure \ref{fig:packet-behavior} compares uniform (i.e. {\footnotesize{CBR}}) packet behavior with the trace of an actual flow. In the uniform model, the cumulative data sent increases smoothly over time, whereas in the actual packet trace, the transmissions are variable both in size and in inter-packet interval, leading to a ``lumpy'' cumulative data plot.}
\label{fig:trace-vs-uniform}
\end{center}
\vspace{-1.25em}
\end{figure*}

@@

\begin{sidewaysfigure*}
%\begin{figure*}[htb]
\subfloat[\textbf{Application:} Average End-to-End Delay]%
{\includegraphics[width=4.45in]{plots/Application/delay.pdf}}
\subfloat[\textbf{Application:} Packet Delivery Ratio]%
{\includegraphics[width=4.45in]{plots/Application/delivery_ratio.pdf}}
\subfloat[\textbf{Application:} Average Received Throughput]%
{\includegraphics[width=4.45in]{plots/Application/throughput.pdf}}
\subfloat[\textbf{Network:} AODV Control Overhead]%
{\includegraphics[width=4.45in]{plots/Network/control_packets.pdf}}
\subfloat[\textbf{Network:} Packets Dropped in Routing Queue]%
{\includegraphics[width=4.45in]{plots/Network/dropped.pdf}}
\subfloat[\textbf{Link:} 802.11 Control Overhead]%
{\includegraphics[width=4.45in]{plots/Link/overhead.pdf}}%
\caption{Box-and-whisker plots of log-ratio error values for all metrics and traffic models. The lower axis indicates the log-ratio, while the upper axis shows raw ratio values. Each box contains the central majority of log-ratio values: the left and right bounds are at the $25^\text{th}$ and $75^\text{th}$ percentiles. The dark middle line indicates the median value, while the diamond marks the mean. The whiskers (dotted lines) extend to the furthest non-outlier values, while the points beyond that are outliers. The notch in the middle of each bar indicates a $95\%$ confidence interval for the true underlying median value; if two notches do not overlap, they are very unlikely to have the same median.}
\label{fig:box-plots}
%\vspace{-1.25em}
%\end{figure*}
\end{sidewaysfigure*}
